EGGs are adhesive!

### 2025-03-04 calco/main.tex
controesempio

\documentclass[a4paper,UKenglish,cleveref,pdftex,thm-restate,numberwithinsect]{lipics-v2021}
\usepackage{stmaryrd}
\usepackage{bbding}
%\Envelope
\newcommand{\cat}[1]{\ensuremath{\mathbf{#1}}}
\usepackage{amsthm,amsmath,amssymb,mathrsfs, dsfont}
\usepackage{hyperref, cleveref} %simboli matematici
\usepackage[all, cmtip]{xy}

@@

\usepackage{tikz-cd}

%String Diagrams
\usepackage{tikz}
\usepackage[draft]{tikzit}
\input{hypergraph.tikzdefs}
\input{hypergraph.tikzstyles}

@@

\begin{abstract}
The use of rewriting-based visual formalisms is on the rise. 
%
In the formal methods community, this is due also to the introduction of adhesive
categories, where most properties of classical approaches to graph transformation, 
such as those on parallelism and confluence, can be rephrased and proved in a general and 
uniform way.
%
E-graphs (EGGs) are a formalism for program optimisation 
via an efficient implementation of equality saturation. 
In short, EGGs can be  defined as (acyclic) term graphs with an additional notion of 
equivalence on nodes that is closed under the operators of the signature.
Instead of replacing the components of a program, the optimisation step 
is performed by adding new components and linking them to 
the existing ones via an equivalence relation, until an optimal program is reached.
%
This work describes EGGs via adhesive categories. 
Besides the benefits in itself of a formal presentation, which renders the 
properties of the data structure precise, the description of the addition of equivalent 
program components using standard graph transformation tools offers the advantages 
of the adhesive framework in modelling, for example, concurrent updates.
%
%Besides 
%the benefits \emph{per se} of a formal presentation, making precise the properties of 
%the data structure, describing the addition of equivalent program components 
%via the standard tools of graph transformation offers the advantages given by the 
%adhesive framework in modelling e.g. concurrent updates.
\end{abstract}

\section{Introduction}
The introduction of \emph{adhesive categories} marked a watershed moment for the algebraic approaches 
to the rewriting of graph-like structures~\cite{lack2005adhesive,ehrig2006fundamentals}.
%
Until then, key results of the approaches on e.g. parallelism and confluence had to be proven 
over and over again for each different formalism at hand, %let it be either xxx or xx, 
despite the obvious similarity of the procedure.
%
Adhesive categories provides such a disparate set of formalisms with a common abstract framework 
where many of these general results could be recast and uniformly proved once and for all.

\vspace{.1cm}
\noindent
\begin{minipage}[l]{.78\linewidth }In short, following the double-pushout (DPO) approach
to graph transformation~\cite{xxx}, a rule is given by two arrows $: K \to L$ and $r: K \rightarrow R$
and its execution requires a match $m: L \to G$ such that the rewriting step from $G$
to $H$ is given by the diagram aside, whose square are pushouts.
  \end{minipage}%
    \hfill
  \begin{minipage}[r]{.20\linewidth }
    \xymatrix@C=.5cm@R=.5cm{
      L \ar[d]_{m}
      & K \ar[r]^r \ar[l]_{l} \ar[d] & R \ar [d] \\
      G & C \ar[r] \ar[l]                    & H
    }
%    \xymatrix@C=.5cm@R=.5cm{
%      L \ar[d]_{m} \ar@{}[dr]|{(1)}
%      & I \ar[r]^r \ar@{>->}[l]_{l} \ar[d] \ar@{}[dr]|{(2)} & R \ar [d] \\
%      G & C \ar@{->}[r] \ar@{>->}[l]                    & H\\
 %     & J \ar@{->}[ul] \ar@{->}[u]|{k} \ar@{->}[ur]
 %   }
  \end{minipage}
\vspace{.1cm}

\noindent
Thus, $L$ and $R$ are the lift- and right-hand side of the rule, respectively, while $K$ witnesses those parts that must 
be present for the rule to be executed, yet they are not affected by the rule itself.
%
If a category is adhesive, and the arrows of the rules are monomorphisms, then the presence of a match ensures that the
two pushouts exist, hence a rewriting step can be performed.
%
The theory of $\mathcal{M}, \mathcal{N}$-adhesivity~\cite{xxx} extend the core framework, still ensuring that if the arrow 
of the rules are in $\mathcal{M}$ and the match is in $\mathcal{N}$, then the classical results from the theory of graph 
transformation can still be lifted~\cite{xxx}. 
%
However, despite the elegance and effectiveness and the obvious advantages of proving a category 
to be adhesive, proving that a given category satisfies the conditions 
for being adhesive can be a daunting task. For this reason,  sufficient criteria have been provided for the core 
framework, e.g. that every elementary topos is adhesive \cite{lack2006toposes}, as well as for the extended one of
$\mathcal{M}, \mathcal{N}$-adhesivity~\cite{CastelnovoGM24}.

\emph{E-graphs} (shortly, EGGs) are an up-and-coming formalism for program optimisation and synthesis via a compact 
representation and efficient implementation of equality saturation. 
%
Albeit a classical data structure~\cite{DetlefsNS05}, EGGs received 
new impulse after the seminal~\cite{WillseyNWFTP21} and
developed a thriving community, as witnessed by the official website~\cite{eggs}.
%
The key idea of rewriting-based program optimisation is to perform the manipulation of a syntactical description 
of a program, replacing some of its components in such a way that the semantics is preserved while 
the computational costs of its actual execution are improved. Instead of directly removing sub-programs, EGGs just add the 
new components and link them to the older ones via the equivalence relation, until an optimal program is 
reached and extracted.

@@

Building on the criterion developed in~\cite{CastelnovoGM24}, the aim of this work is to prove that EGGs form an 
$\mathcal{M}, \mathcal{N}$-adhesive category for a suitable choice of $\mathcal{M}$ and $\mathcal{N}$.
The advantages derived from this characterisation are two-fold. On the one side, 
we put the benefits \emph{per se} of a formal presentation, making precise the properties of the data structure. 
On the other side, describing the optimisation steps via the DPO approach
offers the tools for modelling their parallel and concurrent execution
and for proving their confluence and termination.

\emph{Synopsis}
The paper has the following structure. 
In Section~\ref{sec:ade} we briefly recall 
the theory of $\mathcal{M}$-adhesive categories
and of kernel pairs.
In Section~\ref{sec:hyper} we present the graphical structures of our interest, 
 (labelled) hypergraphs and term graphs, and we provide a
functorial characterisation, which allows for proving their adhesivity properties.
This is expanded in Section~\ref{hypereq} for describing hypergraphs 
and term graphs with equivalence and in Section~\ref{eggs} for capturing
their variants where the equivalences are closed with respect to operator application,
thus subsuming EGGs.
%
In Section~\ref{rewriting} we put the former machinery at work, showing how the optimisation steps
can be rephrased as the application of term graph rewriting rules.
%
Finally, in Section~\ref{conclusioni} we draw our conclusions, hint at future endeavours and offer some 
brief remarks on related works.
%
For the sake of space, the proofs appear in the appendices. 
%For the sake of completeness, 
%and in order to fix the notation, we prove all the results 
%recalled in the background section, besides those that are original 
%to our work.

@@

\begin{restatable}{lemma}{kpp}\label{lemma:kern_pairs_pres_pullbacks}
	Let $f\colon X \to Y$ and $g\colon Z \to W$ be two arrows admitting kernel pairs and suppose that the solid part of the three squares below is given. 
	%
	If the leftmost square is commutative, then there is a a unique arrow $k_h: K_f \to K_g$ making the other two commutative
	\[\xymatrix@R=15pt{X \ar[r]^{h}\ar[d]_{f}& Z \ar[d]^{g} & K_f \ar@{.>}[r]^{k_h} \ar[d]_{\pi^1_f}& K_g \ar[d]^{\pi^1_g} & K_f  \ar@{.>}[r]^{k_h} \ar[d]_{\pi^2_f}& K_g \ar[d]^{\pi^2_g}\\ Y \ar[r]_t& W & X \ar[r]_{h} & Z & X\ar[r]_{h} &Z}\]

		Moreover, it holds
		\begin{enumerate}
			\item if the leftmost square is a pullback, then also the other two are so;
			\item if $h$ is mono then $k_h$ is mono too.
		\end{enumerate} 
\end{restatable}
%[Proof in \Cref{kpp-proof}]

The previous lemma allows us to deduce the following results in an $\mathcal{M}$-adhesive context.
\todo{A questo punto la proposizione seguente possiamo anche buttarla via se serve recuperare spazio}
\noindent
\parbox{7.5cm}{
\begin{restatable}{proposition}{mpo}\label{lem:mpo}
	Let $\X$ be a strict $\mathcal{M}$-adhesive category with all pullbacks, and suppose that in the cube aside the top is an $\mathcal{M}$-pushout and all the vertical faces are pullbacks. Then the right square is a pushout.
\end{restatable}}\hfill 
\parbox{6cm}{\xymatrix@C=10pt@R=6pt{&A'\ar[dd]|\hole_(.65){a}\ar[rr]^{f'} \ar@{>->}[dl]_(.55){m'} && B' \ar[dd]^{b} \ar@{>->}[dl]_(.55){n'} & K_a\ar[rr]^{k_{f'}} \ar[dd]_{k_{m'}}&& K_b \ar[dd]^{k_{n'}} \\ C'  \ar[dd]_{c}\ar[rr]^(.7){g'} & & D' \ar[dd]_(.3){d}\\&A\ar[rr]|\hole^(.65){f} \ar[dl]^{m} && B \ar[dl]^{n}  & K_{c} \ar[rr]_{k_{g'}}&& K_d\\C \ar[rr]_{g} & & D }}

@@

\noindent
\parbox{7.5cm}{
	\begin{restatable}{proposition}{mps}\label{prop:kerset}
		Suppose that in $\Set$ the cube aside is given, whose top and bootom faces are pushout and such that $m\colon A\mto C$ and $m'\colon A'\mto C'4$ are injections. Then the right square, made by the kernel pairs of the vertical arrows is a pushout.
\end{restatable}}\hfill 
\parbox{6cm}{\xymatrix@C=10pt@R=6pt{&A'\ar[dd]|\hole_(.65){a}\ar[rr]^{f'} \ar@{>->}[dl]_(.55){m'} && B' \ar[dd]^{b} \ar@{>->}[dl]_(.55){n'} & K_a\ar[rr]^{k_{f'}} \ar[dd]_{k_{m'}}&& K_b \ar[dd]^{k_{n'}} \\ C'  \ar[dd]_{c}\ar[rr]^(.7){g'} & & D' \ar[dd]_(.3){d}\\&A\ar[rr]|\hole^(.65){f} \ar@{>->}[dl]^{m} && B \ar@{>->}[dl]^{n}  & K_{c} \ar[rr]_{k_{g'}}&& K_d\\C \ar[rr]_{g} & & D }}

@@

We can represent hypergraphs graphically. We use dots to denote nodes and boxex to denote hyperedges, the name of a node or of an hyperedge is put near the corresponding dot or box. Sources and targets are represented by lines between dots and squares: the lines from the sources of an hyperedge comes from the left of the box, while the lines to the targets exit from the right of the box.
%We  decorate the arrow corresponding to the $i^{th}$ letter of a target or a source with a label $i$.
%\todo{Ho lasciato un unico esempio (che è anche un term graph): occorre riscrivere questo paragrafo una volta che abbiamo deciso come rappresentare i grafi}
\iffalse 
\begin{example}Take $V_{\mathcal{G}}$ to be be $\{v_1, v_2, v_3, v_4, v_5\}$ and $E_{\mathcal{G}}$ to be $\{h_1, h_2, h_3\}$. Sources and targets are given by:
	\[\begin{matrix}
		s_{\mathcal{G}}(h_1)\colon 2\to V_{\mathcal{G}}  & \begin{matrix}
			0 \mapsto v_1\\
			1\mapsto v_2
		\end{matrix} && s_{\mathcal{G}}(h_2)\colon 2\to V_{\mathcal{G}} & \begin{matrix}
			0 \mapsto v_3\\
			1\mapsto v_4 
		\end{matrix} && s_{\mathcal{G}}(h_3)\colon 1\to V_{\mathcal{G}} & 
		0 \mapsto v_5\\
		t_{\mathcal{G}}(h_1)\colon 2\to V_{\mathcal{G}} & \begin{matrix}
			0 \mapsto v_3\\
			1\mapsto v_4
		\end{matrix} && t_{\mathcal{G}}(h_2)\colon 2\to V_{\mathcal{G}} & 0\mapsto v_5 && t_{\mathcal{G}}(h_3)\colon 0\to V_{\mathcal{G}} &  t_{\mathcal{G}}(h_3)=?_{ V_{\mathcal{G}}} 
	\end{matrix}\]

@@

			\draw[->-=.5](1.75,-0.9)..controls(1.8,-1.3)and(2.5,-1.5)..(D)node[pos=0.5, below,font=\fontsize{7}{0}\selectfont]{$2$};
			\draw[->-=.5] (1.75,-0.6)..controls(1.8,-0.25)and(2.5,0)..(C) node[pos=0.5, above,font=\fontsize{7}{0}\selectfont]{$1$};
		\end{tikzpicture}
	\end{center}
\end{example}

\begin{example}\label[example]{exa_21} Let $V_{\mathcal{G}}$ be as in the previous example and $E_{\mathcal{G}}=\{h_1, h_2, h_3\}$.	Then we define
	\[\begin{matrix}
		A && s_{\mathcal{G}}(h_1)\colon 0\to V_{\mathcal{G}}  & s_{\mathcal{G}}(h_1)=?_{V_\mathcal{G}} && s_{\mathcal{G}}(h_2)\colon 2\to V_{\mathcal{G}} & \begin{matrix}
			0 \mapsto v_1\\
			1\mapsto v_2
		\end{matrix}&& s_{\mathcal{G}}(h_3)\colon 2\to V_{\mathcal{G}} & \begin{matrix} 
			0 \mapsto v_1\\
			1\mapsto v_4	
		\end{matrix}\\
			B && t_{\mathcal{G}}(h_1)\colon 1\to V_{\mathcal{G}} & 
		0 \mapsto v_1 && 		t_{\mathcal{G}}(h_2)\colon 1\to V_{\mathcal{G}} & 0\mapsto v_3 &&   t_{\mathcal{G}}(h_3)\colon 1\to V_{\mathcal{G}} & 1\mapsto v_5
	\end{matrix}\]

@@

			\draw[->-=.5] (C)..controls(3.5,-0.75)and(4.2,-0.95)..(4.25,-1.35)node[pos=0.5, above,font=\fontsize{7}{0}\selectfont]{$1$};
			\draw[->-=.5] (D)..controls(3.5,-2.25)and(4.2,-2.05)..(4.25,-1.65)node[pos=0.5, below,font=\fontsize{7}{0}\selectfont]{$2$};
		\end{tikzpicture}
	\end{center}
\end{example}
\fi

\begin{example}\label[example]{exa_2}
	Let $V_\mathcal{G} = \{v_1, v_2, v_3, v_4, v_5\}$, and $E_\mathcal{G} = \{h_1, h_2, h_3, h_4\}$. Then we define:
	\[\begin{matrix}
		s_{\mathcal{G}}(h_1)\colon 0\to V_{\mathcal{G}} & s_{\mathcal{G}}(h_1)=?_{V_\mathcal{G}} &&
		s_{\mathcal{G}}(h_2)\colon 0\to V_{\mathcal{G}} & s_{\mathcal{G}}(h_2)=?_{V_\mathcal{G}}\\
		s_{\mathcal{G}}(h_3)\colon 2\to V_{\mathcal{G}} & \begin{matrix} 
					0 \mapsto v_1\\
					1\mapsto v_2
				\end{matrix}&& s_{\mathcal{G}}(h_4)\colon 2 \to V_\mathcal{G} & \begin{matrix}
					0 \mapsto v_3\\
					1\mapsto v_2
				\end{matrix}\\
		t_{\mathcal{G}}(h_1)\colon 1\to V_{\mathcal{G}} & 0 \mapsto v_1 &&
		t_{\mathcal{G}}(h_2)\colon 1\to V_{\mathcal{G}} & 0\mapsto v_3\\
		t_{\mathcal{G}}(h_3)\colon 1\to V_{\mathcal{G}} & 1\mapsto v_3 &&
		t_\mathcal{G}(h_4)\colon 1 \to V_\mathcal{G} & 0 \mapsto v_4
	\end{matrix}\]

	Now, we can depict $\mathcal{G}$ as

	\begin{center}
        \begin{tikzpicture}
                \begin{pgfonlayer}{nodelayer}
                        \node[style=none] (lab1) at (-1.8, 0.8) {$h_1$};
                        \node[style=small box] (h1) at (-1.8, 0.3) {};
                        \node[style=small box] (2) at (-1.8, -0.5) {};
                        \node[style=none] (lab2) at (-1.8, -1.0) {$h_2$};
                        \node[style=none] (h1b) at (-1.8, 0.3) {};
                        \node[style=none] (th2) at (0, 0.3) {};
                        \node[style=none] (bh2) at (0, -0.3) {};
                        \node[style=none] (h2b) at (0, 0) {};
                        \node[style=none] (th3) at (1.8, 0){};
                        \node[style=none] (bh3) at (1.8, -0.6){};
                        \node[style=none] (h3c) at (1.8, -0.3){};
                        \node[style=none] (2c) at (-1.8, -0.5){};
                        \node[style=none] (lab3) at (0, 0.8) {$h_3$};
                        \node[style=medium box] (h2) at (0, 0) {};
                        \node[style=none] (lab4) at (1.8, 0.5) {$h_4$};
                        \node[style=medium box] (h3) at (1.8, -0.3) {};
                        \node[style=none](labv2) at (-0.9, -0.9) {$v_2$};
                        \node[style=node] (v2) at (-0.9, -0.5) {};
                        \node[style=none] (labv1) at (-0.9, 0.7) {$v_1$};
                        \node[style=node] (v1) at (-0.9, 0.3) {};
                        \node[style=none] (labv3) at (0.9, 0.4) {$v_3$};
                        \node[style=node] (v3) at (0.9, 0) {};
                        \node[style=none] (labv4) at (2.7, 0.1) {$v_4$};
                        \node[style=node] (v4) at (2.7, -0.3) {};
                \end{pgfonlayer}
                \begin{pgfonlayer}{edgelayer}
                        \draw (h1b.center) to (th2.center);
                        \draw (2c.center) to (v2);
                        \draw[in=180, out=30] (v2) to (bh2.center);
                        \draw (h2b.center) to (th3.center);
                        \draw (v4) to (h3c.center);
                        %\draw (h3c) to (v4);
                        \draw[in=180, out=-30](v2) to (bh3);
                \end{pgfonlayer}
        \end{tikzpicture}
	\end{center}

@@

\iffalse
\todo{Decidere se tenere questo esempio}

>>>>

\documentclass[a4paper,UKenglish,cleveref,pdftex,thm-restate,numberwithinsect]{lipics-v2021}
\usepackage{stmaryrd}
\usepackage{bbding}
%\Envelope
\newcommand{\cat}[1]{\ensuremath{\mathbf{#1}}}
\usepackage{amsthm,amsmath,amssymb,mathrsfs, dsfont}
\usepackage{tikz-cd}
\usepackage{hyperref, cleveref} %simboli matematici
\usepackage[all, cmtip]{xy}

@@

%\usepackage{tikz-cd}

%String Diagrams
\usepackage{tikz}
%\input{hypergraph.tikzdefs}
%\input{hypergraphs.tikzstyles}

@@

\begin{abstract}
The use of rewriting-based visual formalisms is on the rise. 
%
In the formal methods community, this is due also to the introduction of adhesive
categories, where most properties of classical approaches to graph transformation, 
such as those on parallelism and causality, can be rephrased and proved in a general and 
uniform way.
%
E-graphs (EGGs) are a formalism for program optimisation 
via an efficient implementation of equality saturation. 
In short, EGGs can be  defined as (acyclic) term graphs with an additional notion of 
equivalence on nodes that is closed under the operators of the signature.
Instead of replacing the components of a program, the optimisation step 
is performed by adding new components and linking them to 
the existing ones via an equivalence relation, until an optimal program is reached.
%
This work describes EGGs via adhesive categories. 
Besides the benefits in itself of a formal presentation, which renders the 
properties of the data structure precise, the description of the addition of equivalent 
program components using standard graph transformation tools offers the advantages 
of the adhesive framework in modelling, for example, concurrent updates.
%
%Besides 
%the benefits \emph{per se} of a formal presentation, making precise the properties of 
%the data structure, describing the addition of equivalent program components 
%via the standard tools of graph transformation offers the advantages given by the 
%adhesive framework in modelling e.g. concurrent updates.
\end{abstract}

\section{Introduction}
\todo{A very nice introduction}
\begin{itemize}
\item general on adhesive categories
%\item general on the use of eggs
\end{itemize}

E-graphs (shortly, EGGs) are an up-and-coming formalism for program optimisation and synthesis via a compact 
representation and efficient implementation of equality saturation. 
%
Albeit a classical data structure~\cite{DetlefsNS05}, EGGs received 
new impulse after the seminal~\cite{WillseyNWFTP21} and
developed a thriving community, as witnessed by the official website~\cite{eggs}.
%
The key idea of rewriting-based program optimisation is to perform the manipulation of a syntactical description 
of a program, replacing some of its components in such a way that the semantics is preserved while improving 
e.g. time or space costs of its actual computation. Instead of directly removing sub-programs, EGGs just add the 
new components and link them to the older ones via the equivalence relation, until an optimal program is 
reached and extracted.

@@

The aim of this work is to describe EGGs by means of adhesive categories. The advantages are two-fold. On the one side, 
we put the benefits \emph{per se} of a formal presentation, making precise the properties of the data structure. 
On the other side, describing the addition of equivalent sub-programs 
in terms of DPO transformation offer the possibility of modelling parallel execution of optimisation steps as well as discussing
about confluence and termination.

\emph{Synopsis}
The paper has the following structure. 
In Section~\ref{sec:ade} we briefly recall 
the theory of $\mathcal{M}$-adhesive categories
and of kernel pairs.
In Section~\ref{sec:hyper} we present the graphical structures of our interest, 
 (labelled, acyclic) hypergraphs and term graphs, and we provide a
functorial characterisation, which allows for proving their adhesivity properties.
This is expanded in Section~\ref{hypereq} for describing hypergraphs 
and term graphs with equivalence and in Section~\ref{eggs} for capturing
their variants where the equivalences are closed with respect to operator application,
thus subsuming EGGs.
%
In Section~\ref{rewriting} we put the former machinery at work, showing how the optimisation steps
can be rephrased as suitable applications of term graph rewriting rules.
%
Finally, in Section~\ref{conclusioni} we draw our conclusions, hint at future endeavours and offer some 
brief remarks on related works.
%
For the sake of space, proofs have been moved to the appendices. 
%For the sake of completeness, 
%and in order to fix the notation, we prove all the results 
%recalled in the background section, besides those that are original 
%to our work.

@@

\begin{restatable}{lemma}{kpp}\label{lemma:kern_pairs_pres_pullbacks}
	Let $f\colon X \to Y$ and $g\colon Z \to W$ be two arrows admitting kernel pairs and suppose that the solid part of the three squares below is given.

	If the leftmost square is commutative, then there exists a a unique arrow $k_h: K_f \to K_g$ making the other two commutative.
	\[\xymatrix@R=15pt{X \ar[r]^{h}\ar[d]_{f}& Z \ar[d]^{g} & K_f \ar@{.>}[r]^{k_h} \ar[d]_{\pi^1_f}& K_g \ar[d]^{\pi^1_g} & K_f  \ar@{.>}[r]^{k_h} \ar[d]_{\pi^2_f}& K_g \ar[d]^{\pi^2_g}\\ Y \ar[r]_t& W & X \ar[r]_{h} & Z & X\ar[r]_{h} &Z}\]

		Moreover, the following hold true:
		\begin{enumerate}
			\item if the leftmost squareis a pullback, then also the other two are so;
			\item if $h$ is mono then $k_h$ is mono too.
		\end{enumerate} 
\end{restatable}
%[Proof in \Cref{kpp-proof}]

The previous result allows us to deduce the following lemma in an $\mathcal{M}$-adhesive context.
\todo{A questo punto la proposizione seguente possiamo anche buttarla via se serve recuperare spazio}
\noindent
\parbox{7.5cm}{
\begin{restatable}{proposition}{mpo}\label{lem:mpo}
	Let $\X$ be a strict $\mathcal{M}$-adhesive category with all pullbacks, and suppose that in the cube aside the top is an $\mathcal{M}$-pushout and all the vertical faces are pullbacks. Then the right square is a pushout.
\end{restatable}}\hfill 
\parbox{6cm}{\xymatrix@C=10pt@R=6pt{&A'\ar[dd]|\hole_(.65){a}\ar[rr]^{f'} \ar@{>->}[dl]_(.55){m'} && B' \ar[dd]^{b} \ar@{>->}[dl]_(.55){n'} & K_a\ar[rr]^{k_{f'}} \ar[dd]_{k_{m'}}&& K_b \ar[dd]^{k_{n'}} \\ C'  \ar[dd]_{c}\ar[rr]^(.7){g'} & & D' \ar[dd]_(.3){d}\\&A\ar[rr]|\hole^(.65){f} \ar[dl]^{m} && B \ar[dl]^{n}  & K_{c} \ar[rr]_{k_{g'}}&& K_d\\C \ar[rr]_{g} & & D }}

@@

\todo{'sta cragna che segue è falsa come una moneta da 3 euro: cancellare}
\noindent
\parbox{7.5cm}{
	\begin{restatable}{proposition}{mps}\label{prop:kerset}
		Suppose that in $\Set$ the cube aside is given, whose top and bottom faces are pushout and such that $m\colon A\mto C$, $m'\colon A'\mto C'$ and $g\colon C\mto D$ are injections. Then the right square, made by the kernel pairs of the vertical arrows, is a pushout.
\end{restatable}}\hfill 
\parbox{6cm}{\xymatrix@C=10pt@R=6pt{&A'\ar[dd]|\hole_(.65){a}\ar[rr]^{f'} \ar@{>->}[dl]_(.55){m'} && B' \ar[dd]^{b} \ar@{>->}[dl]_(.55){n'} & K_a\ar[rr]^{k_{f'}} \ar[dd]_{k_{m'}}&& K_b \ar[dd]^{k_{n'}} \\ C'  \ar[dd]_{c}\ar[rr]^(.7){g'} & & D' \ar[dd]_(.3){d}\\&A\ar[rr]|\hole^(.65){f} \ar@{>->}[dl]^{m} && B \ar@{>->}[dl]^{n}  & K_{c} \ar[rr]_{k_{g'}}&& K_d\\C \ar@{>->}[rr]_{g} & & D }}

@@

We can represent hypergraphs graphically. We use dots to denote nodes and squares to denote hyperedges, the name of a node or of an hyperedge is put near the corresponding dot or square. Sources and targets are represented by lines between dots and squares: the lines from the sources of an hyperedge have an arrowhead in the middle pointing towards the hyperedge, while the lines to the targets have arrowheads pointing to the target nodes.  We  decorate the arrow corresponding to the $i^{th}$ letter of a target or a source with a label $i$.
\todo{Ho lasciato un unico esempio (che è anche un term graph): occorre riscrivere questo paragrafo una volta che abbiamo deciso come rappresentare i grafi}
\iffalse 
\begin{example}Take $V_{\mathcal{G}}$ to be be $\{v_1, v_2, v_3, v_4, v_5\}$ and $E_{\mathcal{G}}$ to be $\{h_1, h_2, h_3\}$. Sources and targets are given by:
	\[\begin{matrix}
		s_{\mathcal{G}}(h_1)\colon 2\to V_{\mathcal{G}}  & \begin{matrix}
			0 \mapsto v_1\\
			1\mapsto v_2
		\end{matrix} && s_{\mathcal{G}}(h_2)\colon 2\to V_{\mathcal{G}} & \begin{matrix}
			0 \mapsto v_3\\
			1\mapsto v_4 
		\end{matrix} && s_{\mathcal{G}}(h_3)\colon 1\to V_{\mathcal{G}} & 
		0 \mapsto v_5\\
		t_{\mathcal{G}}(h_1)\colon 2\to V_{\mathcal{G}} & \begin{matrix}
			0 \mapsto v_3\\
			1\mapsto v_4
		\end{matrix} && t_{\mathcal{G}}(h_2)\colon 2\to V_{\mathcal{G}} & 0\mapsto v_5 && t_{\mathcal{G}}(h_3)\colon 0\to V_{\mathcal{G}} &  t_{\mathcal{G}}(h_3)=?_{ V_{\mathcal{G}}} 
	\end{matrix}\]

@@

			\draw[->-=.5](1.75,-0.9)..controls(1.8,-1.3)and(2.5,-1.5)..(D)node[pos=0.5, below,font=\fontsize{7}{0}\selectfont]{$2$};
			\draw[->-=.5] (1.75,-0.6)..controls(1.8,-0.25)and(2.5,0)..(C) node[pos=0.5, above,font=\fontsize{7}{0}\selectfont]{$1$};
		\end{tikzpicture}
	\end{center}
\end{example}
\fi

\begin{example}\label[example]{exa_21} Let $V_{\mathcal{G}}$ be as in the previous example and $E_{\mathcal{G}}=\{h_1, h_2, h_3\}$.	Then we define
	\[\begin{matrix}
		s_{\mathcal{G}}(h_1)\colon 0\to V_{\mathcal{G}}  & s_{\mathcal{G}}(h_1)=?_{V_\mathcal{G}} && s_{\mathcal{G}}(h_2)\colon 2\to V_{\mathcal{G}} & \begin{matrix}
			0 \mapsto v_1\\
			1\mapsto v_2
		\end{matrix}&& s_{\mathcal{G}}(h_3)\colon 2\to V_{\mathcal{G}} & \begin{matrix} 
			0 \mapsto v_1\\
			1\mapsto v_4	
		\end{matrix}\\
		t_{\mathcal{G}}(h_1)\colon 1\to V_{\mathcal{G}} & 
		0 \mapsto v_1 && 		t_{\mathcal{G}}(h_2)\colon 1\to V_{\mathcal{G}} & 0\mapsto v_3 &&   t_{\mathcal{G}}(h_3)\colon 1\to V_{\mathcal{G}} & 1\mapsto v_5
	\end{matrix}\]

@@

			\draw[->-=.5] (C)..controls(3.5,-0.75)and(4.2,-0.95)..(4.25,-1.35)node[pos=0.5, above,font=\fontsize{7}{0}\selectfont]{$1$};
			\draw[->-=.5] (D)..controls(3.5,-2.25)and(4.2,-2.05)..(4.25,-1.65)node[pos=0.5, below,font=\fontsize{7}{0}\selectfont]{$2$};
		\end{tikzpicture}
	\end{center}
\end{example}

\iffalse
%% Questa roba la devo aggiustare sennò non compila %%
\begin{example}\label[example]{exa_2}
	Let $V_\mathcal{G} = \{v_1, v_2, v_3, v_4, v_5\}$, and $E_\mathcal{G} = \{h_1, h_2, h_3, h_4\}$. Then we define:
	\[\begin{matrix}
		s_{\mathcal{G}}(h_1)\colon 0\to V_{\mathcal{G}} & s_{\mathcal{G}}(h_1)=?_{V_\mathcal{G}} &&
		s_{\mathcal{G}}(h_2)\colon 0\to V_{\mathcal{G}} & s_{\mathcal{G}}(h_2)=?_{V_\mathcal{G}} &&
		s_{\mathcal{G}}(h_3)\colon 2\to V_{\mathcal{G}} & \begin{matrix} 
					0 \mapsto v_1\\
					1\mapsto v_4
				\end{matrix}&&
		s_{\mathcal{G}}(h_4)\colon 2 \to V_\mathcal{G} & \begin{matrix}
					0 \mapsto v_3 \\
					1\mapsto v_2 
				\end{matrix}\\

		t_{\mathcal{G}}(h_1)\colon 1\to V_{\mathcal{G}} & 0 \mapsto v_1 &&
		t_{\mathcal{G}}(h_2)\colon 1\to V_{\mathcal{G}} & 0\mapsto v_3 &&
		t_{\mathcal{G}}(h_3)\colon 1\to V_{\mathcal{G}} & 1\mapsto v_5 &&
		t_\mathcal{G}(h_4)\colon 1 \to V_\mathcal{G} & 0 \mapsto v_4
	\end{matrix}\]
\end{example}
\fi

@@

\todo{Decidere se tenere questo esempio}

\begin{example}\label[example]{exa_3} Let $\Sigma=(O_\Sigma, \ari_\Sigma)$ be an algebraic signature in $\Set$. This simply amount to a set of \emph{operations} with an associated natural number, called \emph{arity}. 	For instance let $\Sigma_G$ be the signature of groups, then $\mathcal{G}^{\Sigma_G}$ can be depicted as
	\begin{center}
		\begin{tikzpicture}
			\node[circle,fill=black,inner sep=0pt,minimum size=6pt,label=above:{$v$}] (V) at (0,0) {};
			\node(E)at(-2, 0.4){$e$};
			\node(M)at(0, 2.15){$\cdot$};
			\node(I)at(2, 0.5){$(-)^{-1}$};
			\draw[->-=.5](-1.75,0)--(V)node[pos=0.5, above,font=\fontsize{7}{0}\selectfont]{$1$};
			\draw[->-=.5](V)..controls(-0.5,0.5)and(-0.8,1)..(-0.25,1.6)node[pos=0.5, right,font=\fontsize{7}{0}\selectfont]{$2$};
			\draw[->-=.5](V)..controls(-1,0.6)and(-1,1.1)..(-0.25,1.9)node[pos=0.5, left,font=\fontsize{7}{0}\selectfont]{$1$};
			\draw[->-=.5](0.25,1.75)..controls(0.8,0.8)and(0.5,0.5)..(V)node[pos=0.5, right,font=\fontsize{7}{0}\selectfont]{$1$};
			\draw[->-=.5](V)--(1.75,0)node[pos=0.5, above,font=\fontsize{7}{0}\selectfont]{$1$};
			\draw[->-=.5](2.25,0)..controls(3.5,0)and(2.5,-2)..(V)node[pos=0.5, below,font=\fontsize{7}{0}\selectfont]{$2$};
			\draw[rounded corners] (-2.25, -0.25) rectangle (-1.75, 0.25) {};
			\draw[rounded corners] (-0.25, 1.5) rectangle (0.25, 2) {};
			\draw[rounded corners] (2.25, -0.25) rectangle (1.75, 0.25) {};
		\end{tikzpicture}
	\end{center}
\end{example}
